### 2014-03-11 thesis/src/chapters/chapter-design.tex
Thesis: fixes part 2

What has all of the above-mentioned features is a self-adaptive system that has a holistic approach to an application scaling. As \cite{brun2009engineering} advocates, one of the first architectures of a self-adaptive system was presented in IBM's blueprint for autonomic computing \cite{IBM06}. IBM's autonomic system is widely recognised and respected and hence we decided to leverage it as a foundation of our work. A concept of autonomic system is alluring due to its four crucial attributes:
\begin{itemize}
 \item \emph{self-configuration} - dynamic adaptation to changing environment such as provisioning new application instances, application tuning, traffic delegation to an external provider
  \item \emph{self-healing} - discovering, diagnosing and reacting to environment disruption such as container outage
  \item \emph{self-optimising} - monitoring and tuning resources: migrating containers, offloading traffic, for example
  \item \emph{self-protecting} - detecting and protecting against threats such as Deny-of-Service by provisioning extra resources
\end{itemize}

@@

\section{Overview}
The previous section dealt with all premises that led to a conclusion that the proposed architecture can rely heavily on the concept of an autonomic system. In this section the further elaboration on the solution is presented.
Figure \ref{img:csap-usage-context} showed the context which the architecture applies to. As one can see, Cloud-SAP can be seen as a proxy between the user and cloud instances and as a coordinator of the latter. Although the cloud instance itself consists of a cloud manager and additional resources, in general, it is perfectly legitimate to consider it a resource as well.

@@

The building blocks of the architecture are:
\begin{itemize}
  \item Managed resources
  \item Touchpoints
  \item Touchpoint autonomic managers
  \item Orchestrating autonomic managers
  \item Knowledge source
\end{itemize}
As resources were identified on various levels of abstraction, starting from the application that is to be deployed and ending on a cloud instance that aggregates different applications and services, we can think of the last three layers (managed resources, touchpoints and touchpoint autonomic managers) as a one logical component which can aggregate and manage other components. This recursive definition applies to identified managed resources described in the next section, apart from application as we consider it the most fundamental and indivisible resource.

\subsection{Managed resources}
\emph{Managed resources} form the lowest layer of a stack. When it comes to application provisioning in a cloud environment we can think of the following resources:
\begin{itemize}
  \item \emph{Application} - application itself can be considered resources which can be managed by an external entity, e.g. by restarting/killing it. Interestingly, an application itself can be an autonomic component with an embedded control loop. However, such setting is completely transparent to Cloud-SAP and therefore cannot be managed by it.
  \item \emph{Execution environment of an application} forms an inherent part of an application lifecycle and its proper tuning can greatly influence its performance,
  \item \emph{Container} -- an isolated and controlled environment in which applications are deployed,
  \item \emph{Stack} -- every application has a technology stack associated with it; what is more, if we confine ourselves to map one instance of a container with only one application, we can think of stack as a composition of application (and its execution environment) and a container,
  \item \emph{Service} -- more complex solutions/applications require interactions with various components that are not necessarily implemented using the same technology stack; service denotes a logical entity that consists of a list of stacks that should form one, indivisible component,
  \item \emph{Cloud instance} -- an entity able to deploy \emph{services}
\end{itemize}
What is more, the listed resources are nested within one another and one's ability to work/run smoothly highly depends on the condition and state of the resource it forms part of. This is shown in figure \ref{img:managed-resources-aggregation}.
\begin{figure}[!ht]
  \begin{center}
    \includegraphics[width=\textwidth]{chapter-design/csap-resources-aggregation}
  \end{center}
  \caption{\emph{Managed resources} identified by Cloud-SAP and their aggregation}
  \label{img:managed-resources-aggregation}
\end{figure}

@@

Each autonomic manager implements the control loop which can be split into four consecutive blocks, which share knowledge among one another. The output from one block forms an input to another. However, it is perfectly possible for a block to perform an action for obtaining side effects, e.g. an action can result in producing relevant information contained in log files, which, in turn, comprises a piece of shared knowledge. Each block implements different behaviour based on the abstraction level and resource they operate on (see figure \ref{img:csap-layered-structure}). Nevertheless it is possible to define their generic behaviour, which embraces the four phases: monitoring, analysis, planning and execution.

@@

\subsection{Orchestrating autonomic manager}
While an autonomic manager which manages a single resource may be satisfactory enough in many business cases, there are situations when additional coordination among managers is essential to attain the certain goal. This can result in incorporating autonomic behaviour within an entity in a system-wide scope.

@@

As the exemplary actions can be regarded complex ones, they cannot be associated with actions taken by autonomic managers of only one type. Thus, an orchestrating autonomic manager should coordinate other managers of any type -- an arbitrary mixture of self-healing, self-configuring, self-optimizing and self-protecting managers. Employment of these concepts in Cloud-SAP will be discussed later in detail.

@@

% radek
\section{Autonomic cloud instance manager}
The main responsibility of this manager is the management of a life cycle of resources of a given cloud provider. It could be thought of an autonomic entity which automates the process of managing resources of data centers that forms the services of a cloud instance. The cloud instance manager governs \emph{stacks} of the user-deployed services and its functions include
\begin{inparaenum}[a)]
\item monitoring the state of cloud provider's infrastructure,
\item providing prediction models in terms of resource utilization levels and possible violations of Service Level Agreements,
\item determining the best place (e.g. a physical host) the stack is to be deployed to,
\item ordering a deployment of a stack and
\item migration or eviction of the stacks that do not conform to Quality-of-Service policies provided by the user.
\end{inparaenum}
Additionally, it exposes the manageability interface for taking request from the orchestrating manager.

\subsection{Managed resource}
The resources of a cloud provider can be associated with a data center comprising physical nodes, i.e. clusters, hosts, etc.

@@

Constraint Programming can be another method that can be implemented \cite{nguyen2009autonomic} to tackle this problem. This is done by trying to optimize the global utility function of resource provisioning costs and application-specific requirements and constraints. In effect, the management of infrastructure is done in an automated fashion with the effort to reach the maximum performance of hosted applications, keeping the provider's cost at the minimum level at the same time.

@@

% radek
\section{Autonomic cloud federation manager}
\emph{Autonomic cloud federation manager} is a building block of the proposed architecture that forms the highest layer of its model (see figure \ref{img:csap-layered-structure}). Additionally, it is an element which directly cooperates with a client by capturing and handling their requests.

This manager has two main responsibilities, which include
\begin{inparaenum}[1)]
\item handling deployment of a service requests from the clients and subsequent requests directly applied to that service (e.g. queries about its state), 
\item scaling the previously-deployed service across different cloud vendors.
\end{inparaenum}
The former involves capturing the request from a client, determining the best cloud vendor or vendors for the described service and its deployment. This behaviour is strictly related with \emph{plan} and \emph{execute} functions of an autonomic manager. The latter, on the other hand, involves \emph{monitoring} the dependant autonomic managers that manage in an autonomic fashion cloud instances, \emph{analysing} obtained knowledge and deciding if taking any auto-scaling action is actually necessary to ensure satisfying Quality-of-Service requirements. If such an action is needed, it must be \emph{planned} and \emph{executed}.

@@

\subsection{Managed resource}
As it was stated in the previous sections and can be seen in figure \ref{img:orchestrating-autonomic-manager}, \emph{cloud instance autonomic managers} are managed resources of this manager. Each cloud instance is a deployment environment for a service or a stack, i.e. it is possible to span the service across different clouds by deploying stacks which make up the given service to different providers. Of course it is legitimate to deploy the whole service to only one provider.
\subsubsection*{Touchpoints}
As the cloud federation manager has to gain information about the state of a given cloud instance and be able to order the deployment of a stack or a service, the cloud instance exposes its manageability interface and its two types:
\begin{asparaenum}
\item[\textbf{Sensors}] -- set of attributes which forms the state of a given cloud instance. Attributes should include inter alia overall resources usage, pricing, geographical location. The more thorough elaboration will be held in the section devoted to \emph{plan and execution manager}.
\item[\textbf{Effectors}] -- set of stack/service management operations, such as deploy, migrate or delete.
\end{asparaenum}

@@

The first one is the characteristic of resources available in a cloud -- not only are they geographically distributed, but they can be under different administrative domains and can vary in accessibility. This places a heavy burden on cooperation possibilities across cloud providers, both in a technical and non-technical way as providers need to maintain the high level of trust in order to effectively  deliver its products. Additionally, different entities forms the resources -- starting from individual devices, through virtual machines and whole workstations, ending with clusters and data centers.

@@

  \begin{asparaenum}
  \item[\textbf{Service deployment}] When the user wants to deploy a service on a cloud infrastructure, they have to prepare its detailed description containing technical (stacks) and Quality-of-Service requirements. Additionally, this component should have access to shared knowledge regarding the information about every cloud instance that forms the federation. This should include resource capacity and utilization rates.
  \item[\textbf{Service scaling}] When an application needs to be scaled, there is available additional knowledge that this component should use of -- detailed information about service's performance up to that point, e.g. service resource utilization level (CPU, I/O), response time, its users profile including their geographical location and so on. This knowledge in a form of a prediction model should be passed to the component by the \emph{analyse} function.
  \end{asparaenum}

@@

\subsection{Monitor and Analyse Autonomic Manager}
\subsubsection*{Introduction}
In many cases designers of applications cannot in advance foresee the detailed requirements of resources of their products, e.g. the number of virtual machines comprising it. This is especially true in case of \emph{start-ups} when it is not known whether the product will catch public attention and who exactly matches the target audience. In such circumstances the cloud provider should be able to quickly adapt to new conditions to prevent the service from being unable to work properly, e.g. take into account the geographical location of the application's users and deploy a service in close proximity to them so as to minimize the latency.

@@

\subsubsection*{Prediction models}
One way to divide prediction models into categories is to classify each one as a \emph{reactive} or a \emph{proactive} one. Since in the introductory section there were given reasons why the reactive approach is inefficient, here is a short overview of proactive approaches to the issue.

@@

The knowledge that the component obtains is about current \emph{topology} (which cloud instances make the cloud federation, what is their geographical location, how are they related, etc.) and \emph{policies} (data required to be compared against policies of deployed serviced). What is more, for the system to be able to perform auto-scaling, the gathered information can be classified as \emph{problem determination knowledge} \cite{IBM06}.

@@

In the introductory section a brief overview of some possible techniques and algorithms that can be used in the analyse component was presented. We do not impose a particular method on this component -- it is up to the system complexity and design, which solution would be more suitable to the given needs.

\section{Summary}
Cloud-SAP fills the gap in existing platforms by providing a standardised approach to scaling applications using a fine-grained actions and operating on different platform levels: starting from an execution environment and ending on a cloud federation. It is based on a concept of an autonomic system, providing methodical but flexible foundation for its implementation. What is more, Cloud-SAP leverages core benefits of such a system: promise of a self-configuration, self-optimisation, self-health, self-protection. It is driven by a customer-defined policies and consequently ensures that customer is served with demanded Quality-of-Service and optimised costs.

>>>>

What has all of the above-mentioned features is a self-adaptive system that has a holistic approach to an application scaling. As \cite{brun2009engineering} advocates, one of the first architectures of a self-adaptive system was presented in IBM's blueprint for autonomic computing \cite{IBM06}. IBM's autonomic system is widely recognised and respected model, hence, we decided to leverage it as a foundation of our work. A concept of autonomic system is alluring due to its four crucial attributes:
\begin{itemize}
 \item \emph{self-configuration} - dynamic adaptation to changing environment such as provisioning new application instances, application tuning, traffic delegation to an external provider
  \item \emph{self-healing} - discovering, diagnosing and reacting to environment disruption such as container outage
  \item \emph{self-optimising} - monitoring and tuning resources: migrating containers, offloading traffic, for example
  \item \emph{self-protecting} - detecting and protecting against threats such as Deny-of-Service by provisioning extra resources
\end{itemize}

@@

\section{Overview}
The previous section dealt with all premises that led to a conclusion that the proposed architecture can rely heavily on the concept of an autonomic system. In this section the further elaboration on the solution is presented.
Figure \ref{img:csap-usage-context} showed the context which the architecture applies to. As one can see, Cloud-SAP can be seen as a proxy between the user and cloud instances and as a coordinator of the latter. Although the cloud instance itself consists of a cloud manager and additional resources, in general, it is perfectly legitimate to consider it as a resource as well.

@@

The building blocks of the architecture are:
\begin{itemize}
  \item Managed resources
  \item Touchpoints
  \item Touchpoint autonomic managers
  \item Orchestrating autonomic managers
  \item Knowledge source
\end{itemize}
As resources were identified on various levels of abstraction, starting from the application that is to be deployed and ending on a cloud instance that aggregates different applications and services, the last three layers (managed resources, touchpoints and touchpoint autonomic managers) can be perceived as one logical component which can aggregate and manage other components. This recursive definition applies to identified managed resources described in the next section, apart from application as we consider it the most fundamental and indivisible resource.

\subsection{Managed resources}
\emph{Managed resources} form the lowest layer of a stack. When it comes to application provisioning in a cloud environment we can think of the following resources:
\begin{itemize}
  \item \emph{Application} - application itself can be considered as a resource which can be managed by an external entity, e.g. by restarting/killing it. Interestingly, an application itself can be an autonomic component with an embedded control loop. However, such setting is completely transparent to Cloud-SAP and therefore cannot be managed by it.
  \item \emph{Execution environment of an application} forms an inherent part of an application lifecycle and its proper tuning can greatly influence its performance,
  \item \emph{Container} -- an isolated and controlled environment in which execution environments are deployed,
  \item \emph{Stack} -- every application has a technology stack associated with it; what is more, if we confine ourselves to map one instance of a container with only one application, we can think of stack as a composition of application (and its execution environment) and a container,
  \item \emph{Service} -- more complex solutions/applications require interactions with various components that are not necessarily implemented using the same technology stack; service denotes a logical entity that consists of stacks that forms one, indivisible component,
  \item \emph{Cloud instance} -- an entity able to deploy \emph{services}
\end{itemize}
What is more, the listed resources are nested within one another and one's ability to work/run smoothly highly depends on the condition and state of the resource it forms part of. This is shown in figure \ref{img:managed-resources-aggregation}.
\begin{figure}[!ht]
  \begin{center}
    \includegraphics[width=\textwidth]{chapter-design/csap-resources-aggregation}
  \end{center}
  \caption{\emph{Managed resources} identified by Cloud-SAP and their aggregation}
  \label{img:managed-resources-aggregation}
\end{figure}

@@

Each autonomic manager implements the control loop which can be split into four consecutive blocks, which share knowledge among one another. The output from one block forms an input to another. Each block implements different behaviour based on the abstraction level and resource they operate on (see figure \ref{img:csap-layered-structure}). Nevertheless, it is possible to define their generic behaviour, which embraces the four phases: monitoring, analysis, planning and execution.

@@

\subsection{Orchestrating autonomic manager}
While an autonomic manager which manages a single resource may be satisfactory in many business cases, there are situations when additional coordination among managers is essential to attain the certain goal. This can result in incorporating autonomic behaviour within an entity in a system-wide scope.

@@

As the exemplary actions can be regarded as complex, they cannot be associated with actions taken by autonomic managers of only one type. Thus, an orchestrating autonomic manager should coordinate other managers of any type -- an arbitrary mixture of self-healing, self-configuring, self-optimizing and self-protecting managers. Employment of these concepts in Cloud-SAP will be discussed later in detail.

@@

% radek
\section{Autonomic cloud instance manager}
The main responsibility of this manager is the management of resources' life cycle of a given cloud provider. It could be thought of an autonomic entity which automates the process of managing resources of data centers that forms the services of a cloud instance. The cloud instance manager governs \emph{stacks} of the user-deployed services and its functions include
\begin{inparaenum}[a)]
\item monitoring the state of cloud provider's infrastructure,
\item providing prediction models in terms of resource utilization levels and possible violations of Service Level Agreements,
\item determining the best place (e.g. a physical host) the stack is to be deployed to,
\item ordering a deployment of a stack and
\item migration or eviction of the stacks that do not conform to Quality-of-Service policies provided by the user.
\end{inparaenum}
Additionally, it exposes the manageability interface for taking request from the orchestrating manager.

\subsection{Managed resource}
The resources of a cloud provider associated with a data center comprising physical nodes, i.e. clusters, hosts, etc.

@@

Constraint Programming is another method that can be implemented \cite{nguyen2009autonomic} to tackle this problem. This is done by trying to optimize the global utility function of resource provisioning costs and application-specific requirements and constraints. In effect, the management of infrastructure is done in an automated fashion with the effort to reach the maximum performance of hosted applications, keeping the provider's cost at the minimum level at the same time.

@@

% radek
\section{Autonomic cloud federation manager}
\emph{Autonomic cloud federation manager} is a building block of the proposed architecture that forms the highest layer of its model (see figure \ref{img:csap-layered-structure}). Additionally, it is an element which directly cooperates with a client by capturing and handling his requests.

This manager has two main responsibilities:
\begin{inparaenum}[1)]
\item handling deployment of a service requests from the clients and subsequent requests directly applied to that service (e.g. queries about its state), 
\item scaling the previously-deployed service across different cloud vendors.
\end{inparaenum}
The former involves capturing the request from a client, determining the best cloud vendor or vendors for the described service and its deployment. This behaviour is strictly related with \emph{plan} and \emph{execute} functions of an autonomic manager. The latter, on the other hand, involves \emph{monitoring} the dependant autonomic managers that manage in an autonomic fashion cloud instances, \emph{analysing} obtained knowledge and deciding if taking any auto-scaling action is actually necessary to ensure satisfying Quality-of-Service requirements. If such an action is needed, it must be \emph{planned} and \emph{executed}.

@@

\subsection{Managed resource}
As it was stated in the previous sections and can be seen in figure \ref{img:orchestrating-autonomic-manager}, \emph{cloud instance autonomic managers} are managed resources. Each cloud instance is a deployment environment for a service or a stack, i.e. it is possible to span the service across different clouds by deploying stacks which make up the given service to different providers. Of course it is legitimate to deploy the whole service to only one provider.
\subsubsection*{Touchpoints}
As the cloud federation manager has to gain information about the state of a given cloud instance and be able to order the deployment of a stack or a service, the cloud instance exposes its manageability interface that consists of:
\begin{asparaenum}
\item[\textbf{Sensors}] -- set of attributes which forms the state of a given cloud instance. Attributes should include inter alia overall resources usage, pricing, geographical location. The more thorough elaboration will be held in the section devoted to \emph{plan and execution manager}.
\item[\textbf{Effectors}] -- set of stack/service management operations, such as deploy, migrate or delete.
\end{asparaenum}

@@

The first one is the characteristic of resources available in a cloud -- not only are they geographically distributed, but they can be under different administrative domains and can vary in accessibility. This places a heavy burden on cooperation across cloud providers, both in a technical and non-technical way as providers need to maintain the high level of trust in order to effectively  deliver its products. Additionally, different entities forms the resources -- starting from individual devices, through virtual machines and whole workstations, ending with clusters and data centers.

@@

  \begin{asparaenum}
  \item[\textbf{Service deployment}] When the user wants to deploy a service on a cloud infrastructure, they have to prepare its detailed description containing technical (stacks) and Quality-of-Service requirements. Additionally, this component should have access to shared knowledge regarding the information about every cloud instance that forms the federation. This should include resource capacity and utilization rates.
  \item[\textbf{Service scaling}] When an application needs to be scaled, there is  additional knowledge available that this component should use -- detailed information about service's performance up to that point, e.g. service resource utilization level (CPU, I/O), response time, its users profile including their geographical location and so on. This knowledge in a form of a prediction model should be passed to the component by the \emph{analyse} function.
  \end{asparaenum}

@@

\subsection{Monitor and Analyse Autonomic Manager}
\subsubsection*{Introduction}
In many cases designers of applications cannot in advance foresee the detailed requirements of resources of their products, e.g. the number of virtual machines comprising it. This is especially true in case of \emph{start-ups} when it is not known whether the product will catch public attention and who exactly matches the target audience. In such circumstances the cloud provider should be able to quickly adapt to new conditions to prevent the service from being unable to work properly, e.g. take into account the geographical location of the application's users and deploy a service in proximity to them so as to minimise the latency.

@@

\subsubsection*{Prediction models}
One way to divide prediction models into categories is to classify each one as a \emph{reactive} or a \emph{proactive} one. Since in the introductory section there were given reasons why the reactive approach is inefficient, here is a short overview of proactive approaches.

@@

The knowledge that the component obtains is about current \emph{topology} (which cloud instances make the cloud federation, what is their geographical location, how are they related, etc.) and \emph{policies} (data required to be compared against policies of deployed service). What is more, for the system to be able to perform auto-scaling, the gathered information can be classified as \emph{problem determination knowledge} \cite{IBM06}.

@@

In the introductory section a brief overview of some possible techniques and algorithms that can be used in the analyse component was presented. We do not impose a particular method on this component -- it is up to the system design, which solution would be more suitable to the given needs.

\section{Summary}
Cloud-SAP fills the gap in existing platforms by providing a standardised approach to scaling applications using a fine-grained actions achieved by operating on different platform levels: starting from an execution environment and ending on a cloud federation. It is based on a concept of an autonomic system, providing methodical but flexible foundation for its implementation. What is more, Cloud-SAP leverages core benefits of such system: promise of a self-configuration, self-optimisation, self-health, self-protection. It is driven by a customer-defined policies and consequently ensures that customer is served with demanded Quality-of-Service and optimised costs.

### 2014-03-11 thesis/src/chapters/chapter-evaluation.tex
Thesis: fixes part 2

\subsection*{Hardware and virtual machines configuration}
All tests that were carried out were run on the same hardware and/or used the same virtual machines so for clarity of presentation their parameters are presented not in a description of every test, but here in table \ref{tab:test-deployment-time-common-hardware-configuration}.

@@

\section{Cost of service deployment}
\subsection*{Description}
The aim of this test case is to test the primary use case of the proposed solution -- deployment of a service with the emphasis of client's \textbf{cost}. It should show that the platform chooses the best mapping between the stacks and cloud providers so that the client's pays the \textbf{lowest} possible price.

@@

Because \emph{Carina} is capable of only horizontal scaling and \emph{Cloud-SAP} of both horizontal and vertical, one can question whether the descent in CPU usage can be actually the same in both cases. To answer this question we want to look into the description of the virtual machines comprising the deployed environment. It states that each VM can use up to 30\% power of the CPU of a host. Thus, we can be fairly sure that adding another virtual machine that uses 0.3 CPU capacity of its host is equivalent with changing the CPU usage settings of already deployed VM from 30 to 60\%.
\subsection*{Preconditions}
\subsubsection*{Environment specification}
\begin{asparaenum}
  \item[\textbf{Auto-scaling policy specifications}] As it is shown in figure \ref{fig:auto-scaling-1cp-cpu-usage-function}, we set in both products the threshold values of CPU usage that trigger performing auto-scaling actions to 20 and 60 percent. To apply this setting, the auto-scaling part of service specification looks as follows: in Carina the user has to add the parameters of an auto-scaling policy in a hash that describes the environment under key \emph{:elasticy\_policy}. It is possible to specify the minimal and the maximal number of virtual machines that forms the environment and, what is most important, expressions whose evaluation results in scaling the application. The part responsible for this is shown in listing \ref{lst:auto-scaling-1cp-carina-service-spec}.\lstinputlisting[caption=Carina service specification used for testing auto-scaling with 1 cloud provider, label=lst:auto-scaling-1cp-carina-service-spec]{auto-scaling-1cp-carina-service-spec} In Cloud-SAP it is a matter of setting appropriate arguments to a specific policy. In this case the policy is \emph{threshold\_model} and values are 20 and 60 for lower and upper bound respectively. The auto-scaling part from service specification is shown in listing \ref{lst:auto-scaling-1cp-cloud-sap-service-spec}. \lstinputlisting[caption=Cloud-SAP service specification used for testing auto-scaling with 1 cloud provider, label=lst:auto-scaling-1cp-cloud-sap-service-spec]{auto-scaling-1cp-cloud-sap-service-spec}
  \item[\textbf{OpenNebula/Carina/Cloud-SAP configuration}] Since Carina and Cloud-SAP used two different instances of OpenNebula installed on separate virtual machines, it is essential the configuration be the same on each of them. Listing \ref{lst:auto-scaling-1cp-one-config} shows the most important OpenNebula excerpt from settings file used in this test case -- polling interval specification, which was set to 30 seconds. To ensure that both products can actually use up-to-date data, they should figuate their policies every 30 seconds or longer. In Carina the user cannot specify this value, because it is hard-coded in source code to 60 seconds. In Cloud-SAP, however, this value was set in a configuration file and was equal to 60 seconds as well as in Carina. \lstinputlisting[caption=OpenNebula configuration excerpt -- information manager, label=lst:auto-scaling-1cp-one-config]{auto-scaling-1cp-one-config}. Vertical scaling mechanism in \emph{Cloud-SAP} is implemented in a way that causes exponential growth of CPU resources consumption. When there is a need to perform a scaling action on CPU usage, current value is taken and increased by 30\%.
  \item[\textbf{Deployment diagrams}] Deployment diagrams show the placement of each component in both installations. They are shown in figures \ref{fig:auto-scaling-1cp-carina-deployment-diagram} and \ref{fig:auto-scaling-1cp-cloud-sap-deployment-diagram}.

@@

\subsubsection*{Test case duration}
The test case lasts 25 minutes as the cpu usage is constant and equals 25\% after 21 minutes and 44 seconds.

@@

The first conclusion is that all auto-scaling actions are performed after exceeding the upper threshold value of CPU usage, this is 60 percent in our case what happens about 5 minutes after the deployment of an environment is completed. CPU consumption remains high for about 2 minutes what enables both products to take appropriate, auto-scaling steps. We expect \emph{Carina} to deploy another virtual machine and \emph{Cloud-SAP} to scale vertically. Then it gradually goes down, which simulates that all taken actions have successfully completed. Then, for another several minutes, the cycle recurs, but this time we expect the proposed solution to scale horizontally. \emph{Carina} is expected to behave as in the previous cycle. Once the cycle has completed, after roughly 21 minutes, the value of CPU consumption remains constant at 25 percent and the test is completed.
\subsection*{Results}
  In the first place it is worth discussing the results of \emph{Carina} and \emph{Cloud-SAP} separately and then compare resource consumption of each solution and its influence on cost paid by the end-user.

@@

\subsubsection*{Cost}
In our scenario cost is roughly equivalent with the CPU usage, so the plot of a cost virtually presents the cpu usage. Figure \ref{fig:auto-scaling-1cp-cost-comparison} shows cpu usage by environment when deployed and configured by two competing products.
\begin{figure}[!ht]
  \begin{center}
    \includegraphics{chapter-evaluation/auto-scaling-1cp-cost-comparison}
  \end{center}
  \caption{Auto-scaling - single-provider based: comparison of cost/CPU usage}
  \label{fig:auto-scaling-1cp-cost-comparison}
\end{figure}
If we were to roughly estimate the cost in both solutions it would be equal in Carina:
\begin{equation}
  7\cdot 0.3 + 10\cdot 0.6 + 8\cdot 0.9 = 15.3 \quad \text{[currency unit]}
\end{equation}
and in Cloud-SAP:
\begin{equation}
  7.0166\cdot 0.3 + 2\cdot 0.39 + 2.01667\cdot0.507 + 7.05\cdot 0.659 + 6.91667\cdot 0.959 \approx 15.186483
\end{equation}

\subsection*{Conclusion}
The obtained results shows that the proposed solution has enormous potential in terms of better utilization of available resources of the cloud and reducing cost paid by cloud consumers.

@@

\subsection*{Conclusion}
As one can can notice, deployment of a java stack takes using \emph{Cloud-SAP} lasts longer than when using \emph{Carina} for the same purpose. Moreover, the more instances are in such stack, the greater difference is. This difference in provisioning time is mainly driven by the fact that \emph{Cloud-SAP} uses a great deal of components in comparison to \emph{Carina}, what is a direct consequence of chosen architecture. Specifically, the request from a client is passed to a broker, then the best provider is chosen and finally the deployment using the select provider and its \emph{Applfow} server takes place. Contrary, \emph{Carina} directly operates on \emph{OpenNebula} giving much simpler flow.

Noticeably, deployment time never was a crucial issue for \emph{Cloud-SAP}, hence, this issue is of a little importance. What is actually important for \emph{Cloud-SAP} is deployment cost and service performance. In other words, slightly greater deployment time is a price that is paid to be sure that resources are deployed optimally and with Quality-of-Service guarantee.

>>>>

\subsection*{Hardware and virtual machines configuration}
All tests that were carried out were run on the same hardware and/or used the same virtual machines as presented in table \ref{tab:test-deployment-time-common-hardware-configuration}.

@@

\section{Cost of service deployment}
\subsection*{Description}
The aim of this test case is to test the primary use case of the proposed solution -- deployment of a service with the emphasis on client's \textbf{cost}. It should show that the platform chooses the best mapping between the stacks and cloud providers so that the client's pays the \textbf{lowest} possible price.

@@

Because \emph{Carina} is capable of only horizontal scaling and \emph{Cloud-SAP} of both horizontal and vertical, one can question whether the descent in CPU usage can be actually the same in both cases. To answer this question we want to look into the description of the virtual machines comprising the deployed environment. It states that each VM can use up to 30\% power of a single host's CPU. Thus, we can be fairly sure that adding another virtual machine that uses 0.3 CPU capacity of its host is equivalent with changing the CPU usage settings of already deployed VM from 30 to 60\%.
\subsection*{Preconditions}
\subsubsection*{Environment specification}
\begin{asparaenum}
  \item[\textbf{Auto-scaling policy specifications}] As it is shown in figure \ref{fig:auto-scaling-1cp-cpu-usage-function}, we set in both products the threshold values of CPU usage that trigger performing auto-scaling actions to 20 and 60 percent. To apply this setting, the auto-scaling part of service specification looks as follows: in Carina the user has to add the parameters of an auto-scaling policy in a hash that describes the environment under key \emph{:elasticy\_policy}. It is possible to specify the minimal and the maximal number of virtual machines that forms the environment and, what is most important, expressions whose evaluation results in scaling the application. The part responsible for this is shown in listing \ref{lst:auto-scaling-1cp-carina-service-spec}.\lstinputlisting[caption=Carina service specification used for testing auto-scaling with 1 cloud provider, label=lst:auto-scaling-1cp-carina-service-spec]{auto-scaling-1cp-carina-service-spec} In Cloud-SAP it is a matter of setting appropriate arguments to a specific policy. In this case the policy is \emph{threshold\_model} and values are 20 and 60 for lower and upper bound respectively. The auto-scaling part from service specification is shown in listing \ref{lst:auto-scaling-1cp-cloud-sap-service-spec}. \lstinputlisting[caption=Cloud-SAP service specification used for testing auto-scaling with 1 cloud provider, label=lst:auto-scaling-1cp-cloud-sap-service-spec]{auto-scaling-1cp-cloud-sap-service-spec}
  \item[\textbf{OpenNebula/Carina/Cloud-SAP configuration}] Since Carina and Cloud-SAP used two different instances of OpenNebula installed on separate virtual machines, it is essential the configuration be the same on each of them. Listing \ref{lst:auto-scaling-1cp-one-config} shows the most important OpenNebula excerpt from settings file used in this test case -- polling interval specification, which was set to 30 seconds. To ensure that both products can actually use up-to-date data, they should figure their policies every 30 seconds or longer. In Carina the user cannot specify this value, because it is hard-coded in source code to 60 seconds. In Cloud-SAP, however, this value was set in a configuration file and was equal to 60 seconds as well as in Carina. \lstinputlisting[caption=OpenNebula configuration excerpt -- information manager, label=lst:auto-scaling-1cp-one-config]{auto-scaling-1cp-one-config}. Vertical scaling mechanism in \emph{Cloud-SAP} is implemented in a way that causes exponential growth of CPU resources consumption. When there is a need to perform a scaling action on CPU usage, current value is taken and increased by 30\%.
  \item[\textbf{Deployment diagrams}] Deployment diagrams show the placement of each component in both installations. They are shown in figures \ref{fig:auto-scaling-1cp-carina-deployment-diagram} and \ref{fig:auto-scaling-1cp-cloud-sap-deployment-diagram}.

@@

\subsubsection*{Test case duration}
The test case lasts 25 minutes as the CPU usage is constant and equals 25\% after 21 minutes and 44 seconds.

@@

The first conclusion is that all auto-scaling actions are performed after exceeding the upper threshold value of CPU usage, this is 60 percent in our case what happens about 5 minutes after the deployment of an environment is completed. CPU consumption remains high for about 2 minutes what enables both products to take appropriate, auto-scaling steps. We expect \emph{Carina} to deploy another virtual machine and \emph{Cloud-SAP} to scale vertically. Then it gradually goes down, which simulates that all taken actions have successfully completed. Then, after another several minutes, the cycle recurs, but this time we expect the proposed solution to scale horizontally. \emph{Carina} is expected to behave as in the previous cycle. Once the cycle has completed, after roughly 21 minutes, the value of CPU consumption remains constant at 25 percent and the test is completed.
\subsection*{Results}
  In the first place it is worth discussing the results of \emph{Carina} and \emph{Cloud-SAP} separately and then compare resource consumption of each solution and its influence on cost paid by the end-user.

@@

\subsubsection*{Cost}
In our scenario cost is roughly equivalent with the CPU usage, so the plot of a cost virtually presents the CPU usage. Figure \ref{fig:auto-scaling-1cp-cost-comparison} shows CPU usage by environment when deployed and configured by two competing products.
\begin{figure}[!ht]
  \begin{center}
    \includegraphics{chapter-evaluation/auto-scaling-1cp-cost-comparison}
  \end{center}
  \caption{Auto-scaling - single-provider based: comparison of cost/CPU usage}
  \label{fig:auto-scaling-1cp-cost-comparison}
\end{figure}
If we were to roughly estimate the cost in both solutions it would be equal in Carina:
\begin{equation}
  7\cdot 0.3 + 10\cdot 0.6 + 8\cdot 0.9 = 15.3 \quad \text{[currency unit]}
\end{equation}
and in Cloud-SAP:
\begin{equation}
  7.0166\cdot 0.3 + 2\cdot 0.39 + 2.01667\cdot0.507 + 7.05\cdot 0.659 + 6.91667\cdot 0.959 \approx 15.186483
\end{equation}

\subsection*{Conclusion}
The obtained results shows that the proposed solution has enormous potential in terms of better utilisation of available resources of the cloud and reducing cost paid by cloud consumers.

@@

\subsection*{Conclusion}
As one can can notice, deployment of a java stack using \emph{Cloud-SAP} lasts longer than when using \emph{Carina} for the same purpose. Moreover, the more instances are in such stack, the greater difference is. This difference in provisioning time is mainly driven by the fact that \emph{Cloud-SAP} uses a great deal of components in comparison to \emph{Carina}, what is a direct consequence of chosen architecture. Specifically, the request from a client is passed to a broker, then the best provider is chosen and finally the deployment using the select provider and its \emph{Applfow} server takes place. Contrary, \emph{Carina} directly operates on \emph{OpenNebula} giving much simpler flow.

Noticeably, deployment time never was a crucial issue for \emph{Cloud-SAP}, hence, this matter is of a little importance. What is actually important for \emph{Cloud-SAP} is deployment cost and service performance. In other words, slightly greater deployment time is a price that is paid to be sure that resources are deployed optimally and with Quality-of-Service guarantee.
